---
title: "eds221-day4-notes"
format: html
editor: visual
execute:
    warning: FALSE
---

## Nested for loops

```{r}
file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

## Functions

```{r}
# conceptually 

function_name <- function(bird, dog) {
  pets = bird + dog
  return(pets)
}

```

```{r}
# function in R

exclaim_age <- function(age) {
  print(paste("I am", age, "years old!"))
}

exclaim_age(age = 10)
```

```{r}
# side example
demoFunction <- function(apples, oranges) {
  print(paste("I love", apples, "and", oranges))
}

demoFunction(4, 9)

demoFunction(apples = 4, oranges = 9)

demoFunction(oranges = 10, apples = 3)
```

```{r}
animal_age <- function(animal, age) {
  if (animal == "dog") {
    print(age * 7)
  } else if (animal == "goat") {
    print(age * 4.7)
  }
}

animal_age("dog", 4)
```

# Functions interactive

Make a function that doubles a value

```{r}
double_value <- function(my_dollars) {
  print(2 * my_dollars)
}

double_value(my_dollars = 10.13)
```

```{r}
monthly_income <- c(42.50, 61.75, 89.50)

double_value(monthly_income)
```

Create a function that returns an awesome restaurant name given a cool animal and your favorite food

```{r}
name_restaurant <- function(animal, food) {
  print(paste0(animal, food))
}

name_restaurant(animal = "hawk", food = "pizza")
```

Given temperature in degrees Fahrenheit, write a function that converts it to Celsius

```{r}
convert_f_to_c <- function(f_temp) {
  print((f_temp-32)/1.8)
}

convert_f_to_c(52)
```

Create a function that given inputs of age and hometown, returns "I am \_ years old, and I'm from \_"

```{r}
origin_statement <- function(age, hometown) {
  print(paste("I am", age, "years old, and I'm from", hometown))
}

origin_statement(age = "25", hometown = "Norwalk")
```

```{r}
double_value2 <- function(dollars) {
  dollars * 2
}
double_value2(dollars = 10) * 0.08

class(double_value2(10))
```

Print statements and return statements

Print: "show me", Return: "store this so I can use it later"

# Functions with conditionals

```{r}
calculate_animal_age <- function(species, age_human_yrs) {
  if (species == "dog") {
    age_human_yrs * 7
  } else if (species == "goat") {
    age_human_yrs * 4.7
  }
  else {
    print(paste("Please enter dog or goat"))
  }
}

calculate_animal_age("dog", 10)
calculate_animal_age("goat", 4)
calculate_animal_age("whale", 10)
```

Write a function that, given temperature in Kelvin, if a user selects "metric" to the scale then return degrees Celsius, or if they select "English" then return degrees in Farenheit

```{r}
convert_kelvin <- function(temperature_type, temp_kelvin) {
  if (temperature_type == "metric") {
    print(temp_kelvin - 273.15)
  } else if (temperature_type == "English") {
    print(temp_kelvin - 459.67)
  }
}

convert_kelvin("metric", 30)
```
